---
title: "THE ARTOMETRICS OF JAPANESE ANIMATION"
format:
  html:
    number-sections: false
    theme: none
    css: artometrics.css
    toc: true
    toc-depth: 2
    toc-title: "IN THIS REPORT"
    toc-location: left
    smooth-scroll: true
    embed-resources: true
  pdf:
    number-sections: false
editor:
  markdown:
    wrap: 72
---

```{r setup}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(lubridate)
library(scales)
library(here)
library(ggrepel)
library(ggtext)

# ── Artometrics color palette ──────────────────────────────────────
art_bg        <- "#FAFAF8"
art_cream     <- "#F2F0EB"
art_dark      <- "#1C1C1E"
art_mid       <- "#6B6B6B"
art_highlight <- "#C0392B"
art_secondary <- "#2C3E6B"
art_muted     <- "#D5D5D5"

# ── Artometrics ggplot2 theme ──────────────────────────────────────
theme_artometrics <- function() {
  theme_minimal(base_family = "Helvetica") +
    theme(
      plot.background    = element_rect(fill = art_cream, color = NA),
      panel.background   = element_rect(fill = art_cream, color = NA),
      panel.grid.major   = element_line(color = art_muted, linewidth = 0.4),
      panel.grid.minor   = element_blank(),
      axis.text          = element_text(color = art_mid, size = 9),
      axis.title         = element_text(color = art_dark, size = 10,
                                        face = "bold"),
      plot.title         = element_markdown(color = art_dark, size = 13,
                                        face = "bold", margin = margin(b = 6)),
      plot.subtitle      = element_markdown(color = art_mid, size = 10,
                                        margin = margin(b = 10)),
      plot.caption       = element_text(color = art_mid, size = 8,
                                        hjust = 0,
                                        margin = margin(t = 10)),
      plot.caption.position = "plot",
      plot.margin        = margin(t = 16, r = 16, b = 16, l = 16),
      legend.background  = element_rect(fill = art_cream, color = NA),
      legend.text        = element_text(color = art_mid, size = 8),
      legend.title       = element_text(color = art_dark, size = 9,
                                        face = "bold"),
      strip.text         = element_text(color = art_dark, size = 9,
                                        face = "bold"),
      strip.background   = element_rect(fill = art_cream, color = NA)
    )
}
```

## FAST FACTS

```{r data-setup}
#| echo: false
#| warning: false
#| message: false

# ============================================================
# Anime TidyTuesday (2019-04-23) — SAFE SETUP (no cartesian blowups)
# Creates:
#   1) anime          = title-level table (1 row per animeID)
#   2) anime_genres   = long table (animeID, genre)
#   3) anime_studios  = long table (animeID, studio)
# ============================================================

# ---- read raw ----
raw_df <- readr::read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2019/2019-04-23/raw_anime.csv",
  show_col_types = FALSE
)

# ---- helpers ----
clean_list_col <- function(x) {
  x |>
    as.character() |>
    stringr::str_remove_all("\\[|\\]|'") |>
    stringr::str_split(",") |>
    purrr::map(~ stringr::str_trim(.x)) |>
    purrr::map(~ .x[.x != "" & !is.na(.x)])
}

extract_aired_dates <- function(aired_chr) {
  aired_chr <- as.character(aired_chr)
  start <- stringr::str_extract(aired_chr, "\\d{4}-\\d{2}-\\d{2}")
  aired_wo_start <- stringr::str_remove(aired_chr, "\\d{4}-\\d{2}-\\d{2}")
  end <- stringr::str_extract(aired_wo_start, "\\d{4}-\\d{2}-\\d{2}")
  tibble(
    start_date = lubridate::ymd(start, quiet = TRUE),
    end_date   = lubridate::ymd(end,   quiet = TRUE)
  )
}

# ---- title-level clean table ----
anime <- raw_df %>%
  bind_cols(extract_aired_dates(raw_df$aired)) %>%
  mutate(
    aired_from_year = year(start_date),
    aired_to_year   = year(end_date)
  ) %>%
  filter(rank != 0, popularity != 0) %>%
  distinct(animeID, .keep_all = TRUE)

# ---- tag tables ----
tags_base <- raw_df %>%
  transmute(
    animeID,
    producers_list = clean_list_col(producers),
    genre_list     = clean_list_col(genre),
    studio_list    = clean_list_col(studio)
  )

anime_genres <- tags_base %>%
  select(animeID, genre_list) %>%
  unnest(genre_list, keep_empty = FALSE) %>%
  rename(genre = genre_list) %>%
  distinct(animeID, genre)

anime_studios <- tags_base %>%
  select(animeID, studio_list) %>%
  unnest(studio_list, keep_empty = FALSE) %>%
  rename(studio = studio_list) %>%
  distinct(animeID, studio)

anime_ids <- anime %>% distinct(animeID)
anime_genres  <- anime_genres  %>% semi_join(anime_ids, by = "animeID")
anime_studios <- anime_studios %>% semi_join(anime_ids, by = "animeID")

stopifnot(nrow(anime) == dplyr::n_distinct(anime$animeID))

# ---- fast facts values ----
anime <- anime %>%
  mutate(aired_from_year = lubridate::year(start_date))

n_titles     <- nrow(anime)
year_min     <- min(anime$aired_from_year, na.rm = TRUE)
year_max     <- max(anime$aired_from_year, na.rm = TRUE)
median_score <- median(anime$score, na.rm = TRUE)
```

```{=html}
<style>
  .facts-grid {
    display: grid;
    grid-template-columns: repeat(3, 1fr);
    gap: 16px;
    margin: 24px 0 32px 0;
  }
  .fact-box {
    background: #F2F0EB;
    border-left: 4px solid #C0392B;
    padding: 18px 20px;
    border-radius: 2px;
  }
  .fact-number {
    display: block;
    font-size: 1.75rem;
    font-weight: 700;
    color: #C0392B;
    line-height: 1.1;
    margin-bottom: 6px;
    font-family: Helvetica, sans-serif;
  }
  .fact-label {
    display: block;
    font-size: 0.82rem;
    color: #6B6B6B;
    line-height: 1.45;
    font-family: Helvetica, sans-serif;
  }
  @media (max-width: 700px) {
    .facts-grid { grid-template-columns: 1fr 1fr; }
  }
</style>

<div class="facts-grid">
  <div class="fact-box">
    <span class="fact-number">13,631</span>
    <span class="fact-label">Unique anime titles in the dataset — from silent-era shorts to 2019 streaming originals</span>
  </div>
  <div class="fact-box">
    <span class="fact-number">1917–2019</span>
    <span class="fact-label">Century-scale data window, covering anime from its earliest theatrical roots to the platform era</span>
  </div>
  <div class="fact-box">
    <span class="fact-number">6.38</span>
    <span class="fact-label">Overall median MAL score — the baseline every studio, format, and genre is measured against</span>
  </div>
  <div class="fact-box">
    <span class="fact-number">8.2%</span>
    <span class="fact-label">Share of TV anime scoring 8+ on MAL — the highest hit-rate of any format in the dataset</span>
  </div>
  <div class="fact-box">
    <span class="fact-number">6</span>
    <span class="fact-label">Formats tracked separately: TV, Movie, OVA, Special, ONA, and Music — each behaves differently</span>
  </div>
  <div class="fact-box">
    <span class="fact-number">0.3%</span>
    <span class="fact-label">Share of TV titles scoring 9+ — how rare a masterpiece-level score actually is, even in the best format</span>
  </div>
</div>
```

### A CENTURY OF DATA

This MyAnimeList dataset covers **13,631 unique anime titles** spanning
**1917 to 2019**—closer to a century-scale archive than a "modern
seasonal TV" snapshot. That scope is what makes it worth analyzing as an
industry story, not just a rankings list. It lets you trace anime from
its earliest roots in short films and prewar theater shorts all the way
through the standardized seasonal production machine of the 2010s. At
the same time, the long horizon also means the dataset includes many
niche or archival entries—shorts, early films, obscure OVAs and
ONAs—that behave very differently from the mainstream TV hits most
readers picture when they hear "anime." The key interpretive move
throughout this analysis is separating **catalog reality** (what exists)
from **fan-canon popularity** (what people actually watch and talk
about).

### MORE THAN JUST TV

The format breakdown shows that **TV is the largest single category but
not the majority of the dataset**. TV accounts for **4,260 titles
(31.3%)**, followed by **Movies (2,657; 19.5%)**, **OVAs (2,378;
17.4%)**, **Specials (2,003; 14.7%)**, **ONAs (1,321; 9.7%)**, and
**Music (1,011; 7.4%)**. In other words, most titles in this archive are
**non-TV**—which matters for every chart that follows. When you see a
studio ranking, a genre score, or a popularity curve, you're looking at
a mixed system that includes theatrical releases, direct-to-video OVAs,
broadcast specials, and web originals alongside the TV series that
dominate the cultural conversation. The numbers don't lie about which
format fans think of first; they just remind you that the industry is
bigger and stranger than the seasonal airing calendar suggests.

### WHAT 6.38 ACTUALLY MEANS

On ratings, the dataset baseline is deliberately modest: the overall
**median MAL score is 6.38** (mean **6.29**, standard deviation
**\~1.10**). That number is the calibration point for everything that
follows. When you rank studios or genres later, the real question isn't
"who broke 8"—it's "who can stay **consistently above 6.38** across a
meaningful number of titles?" A show landing at 7.5 is genuinely
performing above the field. A studio holding a median of 7.0 across 100+
titles is doing something structurally different from one that spiked
once. MAL scores skew toward the center of the scale—most titles live in
a middle band between 5 and 7—which means true greatness (8+) is a
relatively thin slice of the total distribution, and anything labeled
"elite" has cleared a real bar.

### THREE WAYS TO BE THE BEST

The "top titles" lists reveal three distinct definitions of "best," and
they don't fully overlap. **Highest-rated** measures prestige. **Most
members** measures reach. **Most favorited** measures intensity—the kind
of devotion that makes someone hit a heart button, not just add a show
to their backlog. The biggest-audience list is dominated by cultural
gateway shows: **Death Note** (the series that arguably got more people
into anime than any other single title in the 2000s), **Attack on
Titan** (which became a global mainstream event by its final arc),
**Sword Art Online** (controversial but massive, essentially defining
the isekai boom), **One Punch Man** (a meme-to-phenomenon pipeline in a
single season), and **Tokyo Ghoul** (peak Crunchyroll-era viral reach).
These are titles that didn't just find an audience—they recruited one.
Meanwhile the top favorites list skews toward the prestige canon:
**Fullmetal Alchemist: Brotherhood**, which held the #1 MAL score for
over a decade and remains the most common answer to "where do I start?";
**Steins;Gate**, the time-travel thriller that built one of the most
devoted fanbases in the medium; and **Hunter x Hunter (2011)**, which
turned a long-running shonen adaptation into a consensus artistic
achievement. Favorites capture something closer to devotion than raw
scale—these are the shows people recommend to their closest friends, not
just anyone who asks.

### LONGEVITY IS NOT GREATNESS

The longest-running title in the dataset is a perfect counterexample to
the idea that more episodes equals more impact. **Lan Mao** (Blue Cat)—a
Chinese children's TV series, not a Japanese production—tops the episode
count at **3,057 episodes**, with a MAL score around **5.64** and a
small member count. It's a useful wrinkle for two reasons. First, it's a
reminder that this dataset includes non-Japanese animation cataloged on
MAL, meaning "anime" here is a broader category than the strict
Japanese-origin definition. Second, and more importantly, it anchors one
of the clearest through-lines in this analysis: **longevity is usually
an industrial scheduling outcome**, not a signal of universal acclaim.
Long-running shows tend to be kids' programming, domestic serials, or
evergreen formats designed to fill time slots reliably—not the
prestige-driven storytelling that generates high scores and devoted
fanbases. The dataset makes this point cleanly: anime has multiple axes
of bigness—**long-running**, **widely watched**, **highly rated**, and
**most beloved**—and the most interesting analysis comes from showing
how those axes diverge.

## ANIME RELEASES PER YEAR BY TYPE

### Output Volume by Format, 1960s–2018

```{r chart1-releases-by-year}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 6.5

type_colors <- c(
  "TV"      = art_secondary,
  "Movie"   = art_highlight,
  "OVA"     = "#E67E22",
  "Special" = "#8E44AD",
  "ONA"     = "#16A085",
  "Music"   = art_mid
)

if (!("aired_from_year" %in% names(anime))) {
  anime <- anime %>% mutate(aired_from_year = lubridate::year(start_date))
}

valid_types <- c("TV", "Movie", "OVA", "Special", "ONA", "Music")

chart1_df <- anime %>%
  filter(!is.na(aired_from_year),
         !is.na(type), type != "",
         type %in% valid_types) %>%
  count(year = aired_from_year, type, name = "n")

total_by_year <- chart1_df %>%
  group_by(year) %>%
  summarise(total = sum(n), .groups = "drop")

start_year <- total_by_year %>%
  filter(total >= 10) %>%
  summarise(min_year = min(year, na.rm = TRUE)) %>%
  pull(min_year)

end_year <- max(chart1_df$year, na.rm = TRUE) - 1

chart1_df <- chart1_df %>%
  filter(year >= start_year, year <= end_year)

smooth_df <- chart1_df %>%
  group_by(type) %>%
  filter(dplyr::n_distinct(year) >= 12) %>%
  ungroup()

p1 <- ggplot(chart1_df, aes(x = year, y = n, color = type)) +
  geom_line(linewidth = 0.9, alpha = 0.65) +
  geom_smooth(
    data = smooth_df,
    se = FALSE,
    method = "loess",
    span = 0.35,
    linewidth = 1.1
  ) +
  facet_wrap(~ type, scales = "free_y", ncol = 3) +
  scale_color_manual(values = type_colors, guide = "none") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6)) +
  scale_y_continuous(labels = scales::comma) +
  theme_artometrics() +
  labs(
    title    = "An industry that industrialized — then diversified",
    subtitle = "Each format tells a different story about how anime scaled. <span style='color:#C0392B'>TV</span> is the engine; ONA is the future arriving.",
    x        = NULL,
    y        = "Titles released",
    caption  = paste0(
      "Data: MyAnimeList via TidyTuesday 2019-04-23 | Window: ",
      start_year, "–", end_year,
      " | LOESS trend shown where sufficient data exists\n— ARTOMETRICS"
    )
  )

p1

type_totals <- chart1_df %>% count(type, wt = n, name = "total_titles")
keep_types  <- type_totals %>% filter(total_titles >= 50) %>% pull(type)
chart1_df   <- chart1_df %>% filter(type %in% keep_types)

ggsave(
  filename = "chart1_releases_by_year.png",
  plot     = p1,
  path     = ".",
  width    = 12, height = 7, dpi = 300, bg = "white"
)
```

### BEFORE THE MACHINE

This chart shows anime transforming from a low-volume, film-anchored
medium into a modern multi-pipeline industry. In the early decades,
**Movies** are the main visible output channel—and even then, output is
sparse and irregular. The 1958 release of *Hakujaden* (Toei Animation's
first color feature film, widely considered the first anime feature in
the modern sense) represents the kind of milestone that exists in a thin
catalog rather than a crowded one. Production in this era is driven by
individual studios making deliberate theatrical bets, not by an
always-on serial system. The long flat baseline across most formats
isn't "no anime existed"—it's that the industry hadn't yet evolved the
infrastructure, the distribution networks, or the financing mechanisms
to ship content at scale. What you're seeing in the early flat lines is
a craft economy, not an industrial one.

### THE HOME VIDEO EXPLOSION

The first real regime shift appears in the **late 1980s into the early
1990s**, where **OVA** output spikes sharply—and the spike is
historically legible once you name the titles driving it. *Megazone 23*
(1985) is often cited as the first major OVA release, a direct-to-video
science fiction film that proved the format could carry serious,
adult-oriented storytelling outside the constraints of TV scheduling.
What followed was a boom: *Bubblegum Crisis* (1987), *Gunbuster* (1988),
*Record of Lodoss War* (1990), *Tenchi Muyo!* (1992)—a wave of
productions that used home video as a distribution model rather than a
workaround. OVAs were direct-to-consumer, less constrained by broadcast
standards, and often used for higher-risk, niche, or premium projects
that TV wouldn't greenlight. The key point is that OVAs aren't just a
format—they're evidence of a **distribution shift**. VHS made a new
production lane economically viable, and the chart shows exactly when
that lane opened. Their gradual decline through the 1990s and 2000s
tracks the same logic in reverse: as TV and later streaming became
dominant channels, the specific niche OVAs filled started to close.

### HOW TV TOOK OVER

The modern industrial takeoff runs from the **late 1990s through the
2010s**, driven by multiple pipelines scaling simultaneously. **TV** is
the core engine, and the inflection point is the mid-to-late 1990s:
*Neon Genesis Evangelion* (1995) didn't just become a phenomenon—it
proved that TV anime could sustain psychological ambition and cultural
impact serious enough to force the industry to treat the format as
prestige territory, not just children's programming. What followed was
the build-out: the standardized seasonal production model (12–13
episodes per cour, committee-based financing, quarterly release cadence)
that turned TV anime from irregular output into a reliable throughput
machine. By the early 2000s, that machine was running the Big Three
simultaneously—*One Piece* (1999–), *Naruto* (2002–2007), *Bleach*
(2004–2012)—each logging hundreds of episodes and anchoring entire
generations of fans. The steepening curve through the 2010s reflects
what happens when that industrial system meets global streaming
distribution: more titles, faster cycles, wider reach, and a structural
incentive to keep the pipeline full.

### THE GLUE BETWEEN SEASONS

**Specials** rise hard in the 2000s–2010s, which fits how the franchise
economy works once TV is the dominant format. Specials aren't standalone
projects—they're how large franchises maintain engagement outside the
strict TV-season cadence. Recap episodes, bonus OVA-style installments
packaged as "specials," promotional one-offs tied to Blu-ray releases,
crossover events, holiday installments—all of these get logged as
Specials in the dataset. As the Big Three and their successors built out
massive franchises with passionate year-round audiences, Specials became
the connective tissue between cours. Think of them as a **flexible
support channel** that expands as the broader franchise ecosystem
expands: the more properties a studio has running, the more
supplementary content gets produced and cataloged alongside them.

### THE PLATFORM ERA ARRIVES

The most important late-era signal is **ONA**. It's essentially
nonexistent for most of the timeline and then explodes in the
**2010s**—a clean structural break marking the arrival of web-first
distribution as a real production lane rather than an exception. Early
ONAs were often short-form experiments: *Shelter* (2016, the Porter
Robinson collaboration that went viral), YouTube-native shorts,
promotional animations for games or music releases. Then the platform
era arrived in force. Netflix's push into anime originals—*Devilman
Crybaby* (2018, Masaaki Yuasa's reimagining that became the template for
what a streaming-native anime could look and feel like), *Violet
Evergarden* (co-produced with Kyoto Animation), *Aggretsuko* (a
workplace comedy short-form series that found a global audience on the
platform)—changed what "web-first" could mean at scale. ONAs reflect a
structural reality: **digital release lowers the barrier to entry,
shortens cycles, diversifies episode structures, and creates a
production lane that doesn't depend on a broadcast slot or a home video
window.**

### EVERYTHING GETS CATALOGED NOW

The **Music** facet ramps late and fast, which reflects two things at
once: real growth in music video and promotional short production tied
to the franchise economy, and a cataloging effect where modern web
ecosystems index small-format releases far more completely than older
archives did. Music videos and short animated promotions have always
existed in the industry—idol tie-ins, OP/ED animation shorts, character
songs—but they're now being logged at scale. Because Music is
lower-count and irregular, it's naturally more volatile year-to-year.
Its late rise still reinforces the same macro thesis as ONA: **the 2010s
aren't just more anime—they're more formats, more release types, and
more of the industry's total output being tracked.**

### INDUSTRIALIZATION, THEN DIVERSIFICATION

The chart's overall shape is an industry history in miniature: a long
flat baseline, a decisive takeoff in the 1990s anchored by TV, a plateau
and then acceleration in the 2000s, and a final diversification phase in
the 2010s where every format except OVA grows simultaneously. That
diversification is where the industry is headed when the dataset ends in
2019. More titles, more platforms, more release types, more markets—and
underneath all of it, **TV still running as the primary engine**, its
line higher and steeper than everything else in the chart.

## HIT-RATE OF HIGH-SCORING TITLES BY FORMAT

### Share of Titles Scoring 8+ by Anime Format

```{r chart2-hitrate-by-format}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 6.5

chart2_df <- anime %>%
  filter(!is.na(score), score > 0,
         !is.na(type), type %in% c("TV","Movie","OVA","Special","ONA","Music")) %>%
  distinct(animeID, .keep_all = TRUE)

type_quality <- chart2_df %>%
  group_by(type) %>%
  summarise(
    n_titles     = n(),
    median_score = median(score, na.rm = TRUE),
    n_ge_8  = sum(score >= 8,   na.rm = TRUE),
    n_ge_85 = sum(score >= 8.5, na.rm = TRUE),
    n_ge_9  = sum(score >= 9,   na.rm = TRUE),
    p_ge_8  = mean(score >= 8,   na.rm = TRUE),
    p_ge_85 = mean(score >= 8.5, na.rm = TRUE),
    p_ge_9  = mean(score >= 9,   na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    type = forcats::fct_reorder(type, p_ge_8),
    label = paste0(
      scales::percent(p_ge_8, accuracy = 0.1),
      " (", n_ge_8, "/", scales::comma(n_titles), ")"
    ),
    is_tv = type == "TV"
  )

p2 <- ggplot(type_quality, aes(x = p_ge_8, y = type,
                                fill = is_tv)) +
  geom_col(alpha = 0.9) +
  geom_text(aes(label = label), hjust = -0.05, size = 3.4,
            color = art_dark) +
  scale_fill_manual(values = c("TRUE" = art_highlight,
                               "FALSE" = art_secondary),
                    guide = "none") +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, max(type_quality$p_ge_8) * 1.22)
  ) +
  theme_artometrics() +
  theme(panel.grid.major.y = element_blank()) +
  labs(
    title    = "<span style='color:#C0392B'>TV</span> is the highest-yield format — by a meaningful margin",
    subtitle = "Share of titles scoring 8+ by format. Labels show count of qualifying titles out of total.",
    x        = "Percent of titles scoring 8+",
    y        = NULL,
    caption  = "Data: MyAnimeList via TidyTuesday 2019-04-23\n— ARTOMETRICS"
  )

p2

ggsave(
  filename = "chart2_hitrate_by_format.png",
  plot     = p2,
  path     = ".",
  width    = 10, height = 6, dpi = 300, bg = "white"
)
```

### YOUR ODDS OF FINDING SOMETHING GREAT

This chart answers a brutally practical viewer question: **"If I pick a
title at random in this format, what are the odds it's actually
great?"** Instead of comparing average scores—which compress differences
and hide the elite tail—you're measuring **hit-rate**: the share of
titles that clear an "I'd recommend this" threshold. Here that threshold
is **8+**, with **8.5+** and **9+** included as progressively rarer
prestige tiers. The first headline is clear: **TV is the highest-yield
hunting ground for 8+ titles**. About **8.2%** of TV titles are rated
**8+** (**346 out of 4,240**), meaningfully higher than every other
format. The next closest is **Movies at 4.8% (128/2,640)**, then **OVA
at 3.0% (70/2,348)**, **Special at 2.1% (41/1,996)**, **ONA at 1.2%
(16/1,308)**, and **Music at 0.2% (2/985)**. Read that as a simple
random-pick expectation: out of 100 picks, TV gives you \~8 hits, Movies
\~5, OVAs \~3, Specials \~2, ONAs \~1, and Music essentially zero.
**It's probability, not vibes.**

### WHY TV WINS

What makes TV's lead especially meaningful is that it's not just "TV has
more titles." You've already normalized by format—TV is winning on
**rate**, which means something structural is happening. TV is where the
medium's consensus classics tend to live: *Fullmetal Alchemist:
Brotherhood* (MAL's long-running #1, a 64-episode TV series), *Attack on
Titan*, *Steins;Gate*, *Hunter x Hunter (2011)*, *Vinland Saga*, *Demon
Slayer*---all TV productions. Weekly release cycles build sustained
communities, long-form storytelling allows character investment that
movies can't match in runtime, and seasonal discourse creates the kind
of collective attention that stabilizes a title's reputation at the top
of the rating scale. TV isn't only the main output pipeline---it's the
**highest-yield engine for top-end audience approval**, not because
every TV show is good, but because the format's conditions favor the
slow-burn consensus that produces durable 8+ scores.

### THE PRESTIGE MYTH

Movies are the interesting second-place result because people often
assume film is anime's prestige lane---and for good reason. *Spirited
Away* won the Academy Award for Best Animated Feature in 2003, the first
(and still only) anime film to do so. *Your Name* (2016) became the
highest-grossing anime film in history at the time of its release,
crossing \$358 million worldwide. *Princess Mononoke*, *Akira*, *Grave
of the Fireflies*---the upper tier of anime film is genuinely among the
most celebrated animation ever made. But the chart reveals what happens
when you look at the whole Movies bucket rather than its famous ceiling:
it's a deeply **mixed market**. Franchise extensions, compilation films
that recut TV arcs for theaters, side stories designed for existing
fans, smaller productions that never find wide distribution---all of
these sit in the same bucket as Ghibli. That mix drags the hit-rate to
4.8%, well below TV's 8.2%, not because anime films are bad, but because
"anime film" as a category includes everything from *Spirited Away* to a
45-minute recap movie packaged as a theatrical event. When a film does
land at the top, it lands hard---but **the odds per random pick are
lower than TV.**

### GEMS IN A DEEP CATALOG

The OVA result is historically coherent. OVAs have produced some of the
medium's most celebrated niche works---*Hellsing Ultimate* (the
gold-standard vampire action OVA), *Legend of the Galactic Heroes* (a
110-episode space opera released entirely as OVAs between 1988 and 1997,
which maintains an exceptionally high MAL score to this day), *FLCL* (6
episodes, endlessly referenced, never quite replicated). These are real
titles that clear the 8+ bar. But as a bucket, OVAs also include
enormous amounts of deep catalog material: forgotten direct-to-video
releases from the 1980s and 1990s, niche genre experiments that found
small audiences, fan-service releases attached to existing franchises.
The hit-rate of **3.0%** reflects that full range---OVAs can absolutely
produce standout titles, but the probability of hitting one by random
selection is lower than TV or Movies. OVAs are a reminder that **format
innovation doesn't automatically translate into mass approval**; it
creates a wider range of outcomes, from cult classics to forgotten
experiments.

### THE SUPPORT LANES

Specials and ONAs function as secondary lanes, and their hit-rates
reflect exactly what those lanes are designed to do. **Specials** exist
to extend franchise attention between seasons---recap episodes, holiday
one-offs, Blu-ray bonus content, promotional shorts. Their job isn't to
produce iconic standalone works; it's to maintain engagement within an
existing property's orbit. A lower hit-rate is logical when you're
pulling from a population of support content rather than ambitious
originals. **ONAs** are more heterogeneous: some are top-tier originals
(*Devilman Crybaby* sits well above 8 and is an ONA), but many are
short-form experiments, platform promotional releases, or web-native
projects where production constraints and distribution priorities differ
fundamentally from TV. The ONA line reads like a platform-era pattern:
**more content, faster cycles, wider variance, and lower probability
that any single title builds the sustained audience consensus that
produces a locked-in 8+ score.**

### HOW RARE IS A MASTERPIECE

The rarity ladder is where the story sharpens further. Even in the
best-performing format, **8.5+ is only 1.9% of TV titles** and **9+ is
only 0.3%**---roughly 11 titles out of 4,240. That's the quiet truth the
chart exposes: **masterpiece-level scores are rare in every format.** TV
doesn't produce floods of 9+ shows; it just gives you the best odds in a
world where truly elite titles are scarce everywhere. The summary table
adds a second context layer: **median score by format**. TV not only has
the best 8+ hit-rate---it has the highest median (**6.79**), meaning the
entire TV distribution is shifted upward relative to other formats.
Movies and OVAs sit at **6.30**, Specials at **6.44**, ONA at **5.69**,
Music at **5.19**. TV isn't winning only because it creates more
hits---it's winning because **the typical TV title is rated higher
too.**

### SHOP NEAR THE ENTRANCE

If you want the best odds of finding an 8+ title quickly, TV is the most
efficient format to browse---not because films and OVAs don't have elite
entries (they do, and some of the most celebrated works in the medium
live there), but because **the hit-rate per random pick is highest in
TV, and the median floor is highest in TV.** Chart 2 pairs directly with
Chart 1: once anime becomes a high-throughput, multi-pipeline industry,
TV remains the lane where audience consensus and top-end ratings
concentrate---even as the platform era expands every other release type
simultaneously.

## POPULARITY RANK VS SCORE

### Do More Popular Anime Score Higher?

```{r chart3-popularity-vs-score}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 6.5

valid_types <- c("TV","Movie","OVA","Special","ONA","Music")

plot_df <- anime %>%
  filter(!is.na(score), score > 0,
         !is.na(popularity), popularity > 0,
         !is.na(type), type %in% valid_types) %>%
  distinct(animeID, .keep_all = TRUE)

pop_cap <- 12000
df_cap <- plot_df %>%
  filter(popularity <= pop_cap) %>%
  mutate(log_pop = log10(popularity))

df_bins <- df_cap %>%
  mutate(
    pop_decile = ntile(log_pop, 10),
    pop_decile = factor(pop_decile, levels = 1:10,
                        labels = paste0("D", 1:10))
  )

chart3_bins <- df_bins %>%
  group_by(pop_decile) %>%
  summarise(
    n        = n(),
    pop_min  = min(popularity, na.rm = TRUE),
    pop_max  = max(popularity, na.rm = TRUE),
    pop_med  = median(popularity, na.rm = TRUE),
    score_med  = median(score, na.rm = TRUE),
    score_q25  = quantile(score, 0.25, na.rm = TRUE),
    score_q75  = quantile(score, 0.75, na.rm = TRUE),
    score_mean = mean(score, na.rm = TRUE),
    share_8p   = mean(score >= 8, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(pop_range = paste0(comma(pop_min), "--", comma(pop_max)))

p3 <- ggplot(chart3_bins, aes(x = pop_decile, y = score_med, group = 1)) +
  geom_ribbon(aes(ymin = score_q25, ymax = score_q75),
              fill = art_secondary, alpha = 0.15) +
  geom_line(linewidth = 1.1, color = art_secondary) +
  geom_point(size = 2.5, color = art_highlight) +
  theme_artometrics() +
  labs(
    title    = "The more popular, the higher the score --- but the gap is structural",
    subtitle = paste0(
      "Titles binned into popularity deciles (D1 = most popular, capped at rank <= ", pop_cap, "). ",
      "Line = median score; band = IQR (25th--75th pct)."
    ),
    x       = "Popularity decile (D1 most popular -> D10 least popular within cap)",
    y       = "MAL score",
    caption = "Data: MyAnimeList via TidyTuesday 2019-04-23\n--- ARTOMETRICS"
  )

p3

ggsave(
  filename = "chart3_popularity_vs_score.png",
  plot     = p3,
  path     = ".",
  width    = 10, height = 6, dpi = 300, bg = "white"
)
```

### THE FRONT GALLERY

Popularity on MyAnimeList is an **attention register**---a rank that
tracks how many people have added a title to their lists, not whether it
is objectively great. But the two turn out to be deeply correlated.
Within the top 12,000 popularity ranks, the **median score slides from
7.73 in D1 (most popular) to 5.75 in D10 (least popular)**---almost a
two-point drop across the distribution. The "elite" zone becomes rare
with startling speed: **34.4% of D1 titles score 8+**, but that
collapses to **12.4% in D2**, **5.5% in D3**, and near-zero through much
of the bottom half. D1 is where the front gallery lives: *Death Note*,
*Fullmetal Alchemist: Brotherhood*, *Attack on Titan*, *Sword Art
Online*, *One Punch Man*---titles with member counts in the millions
that have been watched, finished, rated, argued about, and recommended
so many times that their scores have stabilized into something close to
cultural consensus. The deeper deciles are the archive: vast, uneven,
and mostly unvisited.

### WHY THE POPULAR SHOWS SCORE HIGHER

The correlation isn't accidental---it's structural. When a show like
*Attack on Titan* builds an audience of millions, it gets rated by
millions. That volume stabilizes the score: outlier ratings (the 10s
from devoted fans, the 1s from people who bounced in episode two) get
averaged out by the mass in the middle, and what's left is a reliable
signal. Low-popularity titles have the opposite problem: a few hundred
ratings means a few polarized opinions carry outsized weight, producing
scores that are either artificially inflated or deflated depending on
who showed up. The scores in D1 aren't necessarily "more
correct"---they're **more statistically stable**, which in a rating
ecosystem reads as higher and more trustworthy. Popularity doesn't just
reflect quality; it manufactures the conditions that make quality
legible.

### FAME MAKES ITSELF

There's a second mechanism at work beyond statistical stability:
**visibility creates more visibility**. When *Sword Art Online*
premiered in 2012, it became one of Crunchyroll's most-streamed titles
almost immediately---which meant it appeared at the top of every
recommendations list, every "what to watch next" algorithm, and every
"most popular this season" banner. That placement drove more viewers to
it, which drove more ratings, which kept it ranked high, which drove
more viewers. This is the feedback loop that economists call cumulative
advantage and anime fans just call hype: early traction begets more
traction, and the gap between a show that catches fire in week one and a
show that doesn't can widen dramatically even if the underlying quality
is similar. It's why *Sword Art Online* has more members than many
objectively higher-rated shows---it won the attention game early and the
algorithm kept paying out. The decile gradient you see in this chart is
partly a quality signal and partly a **record of which shows got lucky
with timing, placement, and platform momentum.**

### WHAT LIVES IN THE BACK ROOMS

The bottom deciles aren't full of bad anime---they're full of **unseen
anime**. *Legend of the Galactic Heroes*, a 110-episode space opera OVA
series from 1988, maintains one of the highest MAL scores in the dataset
despite sitting deep in the popularity ranks; it simply never had the
distribution infrastructure to build a mass audience. Dozens of titles
from the 1980s and 1990s exist in similar positions: genuinely
accomplished work that accumulated small, devoted followings and
correspondingly small member counts, leaving their scores statistically
fragile. The wide IQR bands in D7 through D10 tell that story
visually---more variance, more noise, more titles where the score
reflects a few hundred passionate fans rather than a broad cultural
verdict. The back rooms are where discovery happens, but also where the
expected score falls and the signal-to-noise ratio drops. The question
isn't whether great anime exists out there---it does---it's whether
you're willing to accept lower average odds and higher variance in
exchange for the chance of finding something the algorithm never
surfaced.

### THE ALGORITHM PROBLEM

The gradient this chart reveals isn't purely organic---it's partly
manufactured by the discovery systems that sit on top of the catalog.
Streaming platforms and recommendation engines systematically favor
already-popular titles: Crunchyroll's homepage, Netflix's "trending"
row, MAL's own "top anime" lists all amplify visibility for content that
already has high member counts. Research on recommender systems calls
this **popularity bias**---the tendency of algorithms to over-expose the
head of the distribution and under-represent the long tail, compounding
existing visibility gaps over time. In practical terms, this means the
decile gradient you're observing isn't just "popular = good." It's also
"popular = gets recommended = gets more popular = gets more ratings =
looks more authoritative." The back rows of the archive stay
undiscovered not only because the work is less compelling---but because
no system is actively pointing people toward them.

### SHOP WITH YOUR EYES OPEN

The practical takeaway is crisp: **popularity is a strong signal of
higher average ratings, but it is also a mechanism that concentrates
attention.** If you want reliably high-scoring picks, you shop near the
entrance---D1 and D2, where the median score stays above 7 and the 8+
hit-rate is meaningful. If you want discovery, you walk deeper into the
stacks---accepting that the expected score falls, the variance rises,
and the chance of finding something genuinely idiosyncratic and
underseen increases. Neither strategy is wrong. But knowing which one
you're running, and why the distribution looks the way it does, makes
you a smarter browser of a catalog that is much larger and stranger than
any algorithm is going to show you on its own.

## WHICH STUDIOS SCORE HIGHEST?

### Median MAL Score with IQR, Top 20 Studios (Min 20 Titles)

```{r chart4-studio-rankings}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 6.5

studio_df <- anime_studios %>%
  distinct(animeID, studio) %>%
  inner_join(
    anime %>% distinct(animeID, score, type),
    by = "animeID"
  ) %>%
  filter(!is.na(studio), studio != "",
         !is.na(score), score > 0,
         !is.na(type), type != "")

studio_summ <- studio_df %>%
  group_by(studio) %>%
  summarise(
    n_titles     = n(),
    median_score = median(score, na.rm = TRUE),
    q1           = quantile(score, 0.25, na.rm = TRUE),
    q3           = quantile(score, 0.75, na.rm = TRUE),
    tv_share     = mean(type == "TV", na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(n_titles >= 20) %>%
  arrange(desc(median_score)) %>%
  slice_head(n = 20) %>%
  mutate(studio_label = paste0(studio, " (n=", n_titles, ")"))

overall_median <- anime %>%
  distinct(animeID, score) %>%
  summarise(overall_median = median(score, na.rm = TRUE)) %>%
  pull(overall_median)

p4 <- ggplot(studio_summ,
             aes(y = reorder(studio_label, median_score),
                 x = median_score)) +
  geom_errorbarh(aes(xmin = q1, xmax = q3),
                 height = 0, linewidth = 0.7, color = art_mid) +
  geom_point(aes(size = n_titles, color = tv_share), alpha = 0.95) +
  geom_vline(xintercept = overall_median, linetype = "dashed",
             linewidth = 0.7, color = art_highlight, alpha = 0.7) +
  annotate(
    "text",
    x = overall_median,
    y = Inf,
    label = paste0("Overall median = ", round(overall_median, 2)),
    hjust = -0.07, vjust = 1.4,
    size = 3.2, color = art_highlight
  ) +
  scale_color_viridis_c(
    option = "D",
    labels = scales::percent_format(accuracy = 1),
    name   = "TV share"
  ) +
  scale_size_continuous(name = "Titles (n)") +
  coord_cartesian(clip = "off") +
  theme_artometrics() +
  theme(plot.margin = margin(10, 30, 10, 10)) +
  labs(
    title    = "Bones and KyoAni lead --- consistency across volume is the real achievement",
    subtitle = "Top 20 studios by median MAL score (min 20 titles). Bars = IQR. Dot size = output. Color = TV share.",
    x        = "Median MAL score",
    y        = NULL,
    caption  = "Data: MyAnimeList via TidyTuesday 2019-04-23\n--- ARTOMETRICS"
  )

p4

ggsave(
  filename = "chart4_studio_rankings.png",
  plot     = p4,
  path     = ".",
  width    = 12, height = 8, dpi = 300, bg = "white"
)
```

### THE CONSISTENCY QUESTION

This chart answers a deceptively hard question---*which studios score
highest on MyAnimeList*---by using **median score** (the typical title)
instead of averages, which can be hijacked by a single massive hit or a
string of misfires. Across the top-20 studios with at least 20 titles,
the leaders cluster tightly around the **low-to-mid 7s**, while the
overall median of all studio-credited titles sits at **\~6.38**. In
other words: once you filter to studios with real volume, the top tier
isn't separated by a full point---it's separated by **tenths**. This is
exactly what you'd expect in a mature rating ecosystem where most shows
occupy the same band. The leaderboard here isn't about who made the
single best show. It's about **who can stay above the platform's
baseline across a meaningful number of titles**---which is a much harder
thing to do.

### BONES AND KYOANI: THE GOLD STANDARD

At the top, **Bones** and **Kyoto Animation** stand out as the clearest
examples of scale and quality coexisting. Bones' case rests on a catalog
that includes some of the most celebrated titles in the medium:
*Fullmetal Alchemist: Brotherhood*---MAL's long-running #1, a 64-episode
adaptation widely considered the gold standard of shonen
storytelling---alongside *Mob Psycho 100*, *My Hero Academia*, *Soul
Eater*, and *Noragami*. That's not a studio that peaked once; it's a
studio that has produced consensus favorites across multiple genres and
eras. The median stays high because the floor stays high. **Kyoto
Animation** is a different kind of story. Where Bones operates at
volume, KyoAni built its reputation on deliberate, almost obsessive
craft---an in-house training pipeline, refusal to outsource animation
work, and a catalog defined by visual precision: *Clannad*, *K-On!*,
*Hibike! Euphonium*, *A Silent Voice*, *Violet Evergarden*. In July
2019, an arson attack on their main studio killed 36 people---the
deadliest mass killing in Japan in decades. The industry's response was
immediate and global: crowdfunding campaigns, tributes from studios
worldwide, a production community that suddenly understood KyoAni not
just as a company but as a community of artists. That tragedy reframed
how the anime world talks about what studios actually are. Their median
score in this chart is a data point; what sits behind it is something
closer to a legacy.

### THE BOUTIQUE STUDIOS

A second pattern is the presence of smaller studios whose high medians
are built on precision rather than volume. **White Fox** is the clearest
example: a studio whose entire reputation rests on a handful of titles
executed exceptionally well. *Steins;Gate*---the time-travel thriller
that holds one of the highest MAL scores in the dataset and built one of
the most devoted fanbases in the medium---is a White Fox production. So
is *Re:Zero*, which redefined what an isekai could do with psychological
depth and consequence. Two franchises, enormous cultural footprint, very
small catalog. That's the boutique model: high selectivity, high
hit-rate, high sensitivity to any single title underperforming. **David
Production** operates similarly: closely associated with *JoJo's Bizarre
Adventure* (all animated parts) and *Cells at Work*, a studio whose
identity is essentially inseparable from the franchises it's built
around. Both studios show tight IQRs in the chart---small spread around
the median---which is the signature of a studio that doesn't swing
wildly because it doesn't take many swings.

### THE WILD CARD

**Trigger** is the most interesting outlier in this chart because its
wide IQR is not a weakness---it's a brand signature. Founded in 2011 by
ex-Gainax staff including director Hiroyuki Imaishi, Trigger built its
identity on loud, high-style, unapologetically maximalist originals:
*Kill la Kill*, *Little Witch Academia*, *SSSS.Gridman*, *Promare*.
These are shows that don't try to appeal to everyone---they go hard in a
specific direction and trust that the audience who connects with that
direction will connect deeply. The result is exactly what the chart
shows: a high median (the fans who stay love it) alongside a wide IQR
(the audience that bounces bounces hard). When you see a wide IQR bar on
this chart, read it as **variance in reception, not inconsistency in
effort**. Trigger makes exactly the shows it intends to make. Whether
that lands depends entirely on whether you're the kind of viewer they're
making them for.

### THE INDUSTRIAL GIANTS

At the other end of the output spectrum sit studios like **A-1
Pictures** and **Studio Deen**---large dots on the chart, lower medians,
and a catalog range that explains exactly why. A-1 Pictures is part of
the Aniplex/Sony machine, which means it operates at industrial scale:
*Sword Art Online*, *Your Lie in April*, *Anohana*, *Erased*,
*Kaguya-sama: Love Is War* all carry the A-1 credit. That's a range from
consensus classics to broadly popular but divisive titles, and the
median reflects it. You don't run hundreds of productions without
absorbing significant variance in quality---some projects are
passion-driven, others are committee-driven, and the catalog average is
the average of both. **Studio Deen** carries the most historically
loaded reputation in this chart: "Deen quality" became an actual meme in
the anime community, shorthand for inconsistent or underwhelming
animation on otherwise solid source material (the original *Fate/stay
night* adaptation is the canonical example). Their lower median isn't
surprising---it's a data-driven confirmation of a reputation that fans
had already reached on their own.

### WHAT THE DOT SIZE TELLS YOU

The dot size in this chart encodes output volume, and it's worth reading
alongside the median rather than separately. A small dot with a high
median (White Fox, David Production) says: *high precision, low sample
size, sensitive to change*. One underperforming franchise could
meaningfully shift that median. A large dot with a high median (Bones)
says: *the quality is load-bearing across real volume*, which is a
structurally harder thing to sustain. A large dot with a lower median
(A-1 Pictures, Studio Deen) says: *industrial scale with mixed
results*---which is essentially the expected outcome when a studio is
producing enough content to serve every corner of the market. The color
layer---TV share---adds one more read: studios with high TV share are
building their reputation through the seasonal pipeline, while studios
with lower TV share are mixing in films, OVAs, and specials that behave
differently in the rating ecosystem.

### THE REAL WINNER HERE

The studios that top this chart aren't the ones with the single biggest
hit---almost every studio in anime has at least one title that broke
through. The winners are the ones whose **typical show stays above the
platform's baseline despite real volume**. That's a different kind of
achievement than a peak. It requires consistent sourcing (choosing the
right projects), consistent execution (delivering on what the source
material promises), and consistent audience trust (building the kind of
reputation that makes fans show up for your next title before they know
what it is). Bones and Kyoto Animation have all three. That's why
they're at the top of this chart, and why the distance between them and
the industrial giants below them---measured in tenths of a score
point---represents something much larger than the numbers suggest.

## GENRE MAP: POPULARITY VS WELL-RATED

### Median Score vs Median Popularity Rank by Genre Tag

```{r chart5-genre-map}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 6.5

valid_types <- c("TV","Movie","OVA","Special","ONA","Music")

base_titles <- anime %>%
  filter(!is.na(animeID),
         !is.na(score), score > 0,
         !is.na(popularity), popularity > 0,
         !is.na(type), type %in% valid_types) %>%
  distinct(animeID, .keep_all = TRUE)

total_titles <- base_titles %>% distinct(animeID) %>% nrow()

genre_title <- anime_genres %>%
  distinct(animeID, genre) %>%
  filter(!is.na(genre), genre != "") %>%
  inner_join(base_titles %>% select(animeID, score, popularity),
             by = "animeID") %>%
  mutate(log_pop = log10(popularity))

min_n <- 80

genre_stats <- genre_title %>%
  group_by(genre) %>%
  summarise(
    n_titles    = n_distinct(animeID),
    pct_titles  = n_distinct(animeID) / total_titles,
    score_med   = median(score, na.rm = TRUE),
    score_mean  = mean(score, na.rm = TRUE),
    pop_med     = median(popularity, na.rm = TRUE),
    log_pop_med = median(log_pop, na.rm = TRUE),
    share_8p    = mean(score >= 8, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(n_titles >= min_n) %>%
  mutate(hidden_gem = scale(score_med)[,1] + scale(log_pop_med)[,1])

labels_df <- bind_rows(
  genre_stats %>% slice_max(n_titles,  n = 8),
  genre_stats %>% slice_max(score_med, n = 8),
  genre_stats %>% slice_max(hidden_gem, n = 8)
) %>%
  distinct(genre, .keep_all = TRUE)

x_cut <- median(genre_stats$pop_med, na.rm = TRUE)
y_cut <- median(genre_stats$score_med, na.rm = TRUE)

p5 <- ggplot(genre_stats,
             aes(x = pop_med, y = score_med, size = n_titles)) +
  geom_vline(xintercept = x_cut, linetype = "dashed",
             color = art_mid, alpha = 0.4) +
  geom_hline(yintercept = y_cut, linetype = "dashed",
             color = art_mid, alpha = 0.4) +
  geom_point(color = art_secondary, alpha = 0.6) +
  scale_x_reverse(trans = "log10", labels = comma) +
  scale_size_continuous(labels = comma) +
  ggrepel::geom_text_repel(
    data = labels_df,
    aes(label = genre),
    size = 3.2,
    color = art_dark,
    max.overlaps = Inf,
    box.padding = 0.25,
    point.padding = 0.15,
    seed = 42,
    show.legend = FALSE
  ) +
  theme_artometrics() +
  labs(
    title    = "Genre map: what's popular vs what's well-rated?",
    subtitle = paste0(
      "Each point = a genre tag (multi-label). X = median MAL popularity rank (log, reversed). ",
      "Y = median score. Size = titles tagged."
    ),
    x       = "Median popularity rank (log scale; 1 = most popular)",
    y       = "Median MAL score",
    size    = "Titles tagged",
    caption = "Data: MyAnimeList via TidyTuesday 2019-04-23\n--- ARTOMETRICS"
  )

p5

ggsave(
  filename = "chart5_genre_map.png",
  plot     = p5,
  path     = ".",
  width    = 12, height = 8, dpi = 300, bg = "white"
)
```

### THE MAIN HALL

In this chart, each point is a **genre tag** (multi-label, so a single
title can contribute to several bubbles). The x-axis is **median MAL
popularity rank** (lower number = more popular), while the y-axis is
**median MAL score**---a rough map of *attention* versus *reputation*.
The dense middle cluster---**Comedy, Action, Sci-Fi, Fantasy,
Adventure**---is the museum's main hall: large bubbles, broad reach, and
median scores that settle in the mid-6s rather than the top of the
scale. These are the genres that define anime's public face. Action is
where the Big Three live (*One Piece*, *Naruto*, *Bleach*), each running
hundreds of episodes and accumulating members by the million. Comedy is
where *Gintama* sits---a show that somehow managed to be simultaneously
one of the longest-running comedies and one of the highest-rated anime
on the platform, a genuine outlier in a genre where most titles land in
the mid-pack. Sci-Fi has *Steins;Gate* and *Neon Genesis Evangelion* at
the top and hundreds of more obscure titles dragging the median down. In
large genres, **the ceiling is real but so is the floor**, and the
median reflects the full distribution, not just the titles you can name.

### MORE TITLES, MIDDLE SCORES

The pattern across the main hall genres is consistent: **high volume
produces middle medians**. When a genre is a production workhorse---when
studios are shipping Action or Fantasy titles every season because
that's what the market demands---you get a wide range of quality. Some
projects are passion-driven adaptations of beloved source material;
others are committee-greenlit productions chasing a trend. The median
score settles where the full range averages out, which in a competitive,
high-throughput genre is reliably "watchable but not exceptional." This
isn't a knock on these genres---it's what happens when any category
scales. The more Action anime gets made, the more dilution happens
around the hits. *Dragon Ball Z* and *Attack on Titan* don't stop being
great because there are five hundred other action shows in the dataset;
they just stop being statistically representative of what a random
Action pick looks like.

### THE CURATOR'S ROOMS

The more interesting story is in the genres that score like prestige
niches while remaining only moderately popular. **Mystery** stands out
with a **median score of 7.27** and a **16.5% share of titles scoring
8+**---the highest 8+ share of any genre in the table. The titles
driving that number are exactly what you'd expect from a genre that
rewards patience and craft: *Monster* (Naoki Urasawa's slow-burn
thriller, 74 episodes, universally regarded as one of the great anime of
the 2000s), *Hyouka* (Kyoto Animation's understated high school mystery
series, beloved by the fans who find it and largely invisible to
everyone else). **Police** follows a similar logic: *Psycho-Pass* and
*Ghost in the Shell: Stand Alone Complex* anchor a genre that produces
fewer titles but lands them at a higher rate. These are the curator's
rooms---quieter, less trafficked, but reliably satisfying for the
visitor who makes it inside. The genres themselves act as a filter: they
attract viewers who specifically want that kind of storytelling, which
means the audience rating them tends to be self-selected and engaged
rather than casual.

### THE HISTORICAL PROBLEM

The biggest "hidden gem" signal in the dataset belongs to
**Historical**---not because it's the highest-scoring genre, but because
of how far its popularity rank sits from its score. With **964 tagged
titles**, a **median score of 6.56**, and a **median popularity rank of
9,220**, Historical sits deeper in the archive than almost any other
genre while maintaining a respectable reputation among those who find
it. The titles in this space are genuinely accomplished: *Vinland Saga*
(the Viking epic that became one of the most critically praised anime of
the late 2010s), *Kingdom* (one of the longest-running and most
consistently rated historical epics in the dataset), *The Rose of
Versailles* (a foundational shojo title from 1979 that still maintains a
devoted following). What keeps these shows out of the mainstream isn't
quality---it's friction. Historical anime often demands patience, prior
investment in setting and context, and tolerance for slower pacing than
the action-driven seasonal titles that dominate the platform's front
page. In a market where attention is the scarcest resource, friction is
a serious disadvantage regardless of what's on the other side of it.

### WHY KIDS SCORES LAST

**Kids** is the genre where the gap between cultural footprint and MAL
score is most stark. With **2,158 tagged titles (16.0% of the
dataset)**, it's one of the largest genre categories---and yet it shows
a **median score of 5.86** and a **1.0% share of 8+ titles**, the lowest
of any major genre. This isn't because *Doraemon* or *Pokemon* are bad
shows---both are enormous cultural institutions in Japan and globally,
shows that have shaped the childhoods of multiple generations. The low
scores reflect something about **who rates on MAL and what they're
rating for**. MAL's rating culture skews toward adult viewers applying
adult critical frameworks: narrative complexity, thematic depth,
emotional ambiguity, animation craft in service of mature storytelling.
Children's programming is optimized for entirely different
goals---accessibility, repetition, gentle pacing, reassurance---and
those goals are genuinely less legible to a rating system built around
the preferences of a different audience. Kids anime doesn't score low
because it fails. It scores low because it's being judged by a rubric it
was never designed to pass.

### WHAT THE MAP IS REALLY SHOWING

Chart 5 is ultimately a map of how **attention and reputation distribute
unevenly across a large catalog**---and why that unevenness isn't
random. The main hall genres are popular because they're accessible,
high-volume, and platform-promoted; their middle medians reflect the
cost of that scale. The curator's rooms score higher because
self-selected audiences rate more generously and the genre itself acts
as a quality filter; their lower popularity reflects the friction of
asking casual viewers to find them. Historical sits in the archive
because the discovery systems don't surface it, not because the work
doesn't merit attention. Kids scores last because the rating culture
doesn't speak its language. None of this is a verdict on taste---it's a
tour through an attention economy where the map of what's watched and
the map of what's good only partially overlap, and knowing where they
diverge is more useful than pretending they're the same thing.

## LIMITATIONS

### WHAT THIS DATA CAN'T TELL YOU

Every analysis is bounded by its data, and this one is no exception. The
five charts above describe a real industry across a real timespan---but
they describe it from a specific vantage point: a 2019 snapshot of one
platform's user-generated ratings, filtered to titles with nonzero rank
and popularity. Before drawing hard conclusions, it's worth naming the
edges of that vantage point explicitly. Four limitations are significant
enough to change how you should read the numbers.

### THE CUTOFF PROBLEM

This dataset was scraped in 2019, which means everything that has
happened in anime since then is absent. *Demon Slayer: Kimetsu no Yaiba*
broke domestic box office records in Japan and became one of the
fastest-spreading anime properties globally---it's not here. *Jujutsu
Kaisen*, *Chainsaw Man*, *Spy x Family*, the final arc of *Attack on
Titan*---none of it is in this analysis. The studio rankings, genre
medians, and format trends are snapshots of an industry through 2019,
not the industry as it exists today. The structural patterns
described---TV's dominance, the ONA surge, the platform era's
arrival---are real and durable, but the specific numbers would shift
with five more years of data, and some studios' positions on the
leaderboard would shift significantly.

### WHO'S ACTUALLY RATING

MyAnimeList is not a representative sample of all anime viewers---it's a
self-selected community of dedicated fans who track and rate their
watching. Casual viewers, children, people watching dubbed versions on
mainstream platforms, and audiences in markets where MAL is less
dominant (China, Southeast Asia, Latin America) are systematically
underrepresented in the rating data. This matters most for two chart
interpretations: Kids anime scores low partly because children don't
rate on MAL, and popularity ranks reflect the MAL community's
preferences rather than true global viewership. A show like *Doraemon*
that is genuinely watched by millions of Japanese children every week
will appear less "popular" in this dataset than a niche action series
with a passionate MAL-active fanbase.

### THE MULTI-LABEL PROBLEM

Genre tags in this dataset are multi-label, meaning a single title
contributes to every genre it's tagged under. *Fullmetal Alchemist:
Brotherhood* appears in Action, Adventure, Drama, Fantasy, Magic,
Military, and Shounen simultaneously. This inflates the score medians of
genres that tend to appear alongside high-rated titles---genres like
Military or Supernatural may score well partly because they frequently
co-tag with acclaimed Action or Drama series rather than because
standalone Military or Supernatural titles are independently
exceptional. Genre statistics here describe tag behavior as much as they
describe genre quality, and the two are not the same thing.

### SMALL SAMPLES, UNSTABLE MEDIANS

Studio and genre medians become less reliable as sample size shrinks. A
studio like White Fox with a small catalog has a median that could shift
meaningfully if one or two underperforming titles were added. The
minimum thresholds applied in this analysis (20 titles for studios, 80
tagged titles for genres) reduce but don't eliminate this instability.
Boutique studios with high medians should be interpreted as "has
maintained quality across a small but real catalog," not as "will
definitely maintain this median as their catalog grows." Similarly,
genres near the minimum threshold should be treated as directional
signals rather than precise measurements.

## CONCLUSION

### THE INDUSTRY IN FULL

What this dataset describes, across all five charts, is an industry that
transformed itself structurally at least three times in a hundred
years---and is still transforming. It started as a craft economy of
theatrical films and shorts, became a home-video-driven OVA market in
the late 1980s, industrialized into a high-throughput seasonal TV
machine through the 2000s, and entered a platform era in the 2010s where
the number of release formats multiplied alongside the number of titles.
Each of those shifts left a signature in the data: the early Movie
baseline, the OVA spike, the TV acceleration, the ONA explosion. The
charts don't just describe anime---they describe what happens to a
creative industry when it scales.

### THREE THINGS THAT HELD TRUE ACROSS ALL OF IT

Three patterns run through every chart in this analysis and are worth
naming directly. First: **TV is the primary engine**---for output
volume, for audience size, for score concentration, and for the
consensus classics that define how most people think about anime. Every
other format is interesting and sometimes excellent, but none of them
dislodged TV as the format where the industry's most important work
tends to happen. Second: **popularity and quality are correlated but not
identical**. The decile analysis shows they move together---more popular
titles genuinely do score higher on average---but the genre map and the
studio chart both show cases where quality exists without visibility,
and visibility without equivalent quality. The gap between those two
maps is where the most interesting recommendations live. Third:
**consistency beats peaks**. The studios at the top of the ranking chart
aren't necessarily the ones with the single most celebrated title---
they're the ones whose typical show stays above the baseline across real
volume. That's a harder thing to sustain than a single hit, and the data
rewards it accordingly.

### WHERE THIS LEAVES THE READER

If there's a practical takeaway from a century of anime data, it's this:
**the catalog is vast, the attention economy is real, and the best
titles are not evenly distributed across it**. TV and the most popular
deciles give you the best odds per random pick. Mystery, Police, and
Historical give you a higher hit-rate per title if you're willing to
browse outside the front page. Bones and Kyoto Animation give you the
most reliable studio signal if you're using studio as a filter. And if
you go deep enough into the archive---past the platform recommendations,
past the seasonal buzz, past the titles that won the attention game
early---you'll find work that never found its audience, sitting with a
stable score and a small member count, waiting. The dataset ends in
2019. The catalog keeps growing.

## REFERENCES

-   Anderson, C. (2004, October). The long tail. *Wired*.
    https://www.wired.com/2004/10/tail/

-   Hughes, R. (2023). *MyAnimeList top animes* \[Dataset\]. Kaggle.
    https://www.kaggle.com/datasets/reyadgh/my-anime-list-scraped-dataset

-   Mock, T. (2019, April 23). *Anime data* \[TidyTuesday dataset\].
    GitHub.
    https://github.com/rfordatascience/tidytuesday/blob/master/data/2019/2019-04-23/readme.md

-   Rosen, S. (1981). The economics of superstars. *The American
    Economic Review*, *71*(5), 845--858.
    https://home.uchicago.edu/\~vlima/courses/econ201/Superstars.pdf

-   Salganik, M. J., Dodds, P. S., & Watts, D. J. (2006). Experimental
    study of inequality and unpredictability in an artificial cultural
    market. *Science*, *311*(5762), 854--856.

-   Simon, H. A. (1971). Designing organizations for an information-rich
    world. In M. Greenberger (Ed.), *Computers, communications, and the
    public interest* (pp. 37--72). Johns Hopkins Press.

-   Varian, H. R. (1995). The information economy. *Scientific
    American*.
    https://people.ischool.berkeley.edu/\~hal/pages/sciam.html

-   Wikipedia contributors. (2024a). *Bones (studio)*. Wikipedia.
    https://en.wikipedia.org/wiki/Bones\_%28studio%29

-   Wikipedia contributors. (2024b). *Kyoto Animation*. Wikipedia.
    https://en.wikipedia.org/wiki/Kyoto_Animation

-   Wikipedia contributors. (2024c). *Studio Trigger*. Wikipedia.
    https://en.wikipedia.org/wiki/Studio_Trigger

-   Wikipedia contributors. (2024d). *White Fox*. Wikipedia.
    https://en.wikipedia.org/wiki/White_Fox

-   Yin, H., Wang, Q., Zheng, K., Li, Z., & Yang, X. (2023). A survey on
    popularity bias in recommender systems. *arXiv*.
    https://arxiv.org/abs/2308.01118

## EDITOR'S NOTE

*This report was researched, written, designed, and produced in active
collaboration with Claude AI (Anthropic). The data pipeline, statistical
analysis, chart design, written analysis, narrative structure, and
visual styling were all developed through a directed partnership between
human editorial judgment and AI execution. Artometrics was built on the
premise that rigorous analysis and honest process are not in conflict.
The research questions, editorial instincts, interpretive framing, and
brand vision are ours. The execution---every line of R code, every
paragraph of analysis, every design decision---was a collaboration. We
document this not as a disclaimer but as a description of how we
actually work, and as a position: we believe this is what serious data
journalism looks like when the tools available are used honestly and at
full capacity.*

*--- Artometrics Editorial*
