---
title: "ANIME"
subtitle: "The Artometrics of Japanese Animation"
format:
  html:
    number-sections: false
    theme: none
    css: artometrics.css
    toc: true
    toc-depth: 2
    toc-title: "IN THIS REPORT"
    toc-location: left
    smooth-scroll: true
    embed-resources: true
    code-fold: true
    code-tools: true
    include-in-header: art-head.html
    include-before-body: art-body.html
editor:
  markdown:
    wrap: 72
---

```{r setup}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(lubridate)
library(scales)
library(ggrepel)
library(ggtext)
library(ggridges)

dir.create("charts", showWarnings = FALSE)

# ── Artometrics color palette ──────────────────────────────────────
art_bg        <- "#FAFAF8"
art_cream     <- "#F2F0EB"
art_dark      <- "#1C1C1E"
art_mid       <- "#6B6B6B"
art_highlight <- "#C0392B"
art_secondary <- "#2C3E6B"
art_muted     <- "#D5D5D5"

# ── Artometrics ggplot2 theme ──────────────────────────────────────
theme_artometrics <- function() {
  theme_minimal(base_family = "Helvetica") +
    theme(
      plot.background    = element_rect(fill = art_cream, color = NA),
      panel.background   = element_rect(fill = art_cream, color = NA),
      panel.grid.major   = element_line(color = art_muted, linewidth = 0.4),
      panel.grid.minor   = element_blank(),
      axis.text          = element_text(color = art_mid, size = 9),
      axis.title         = element_text(color = art_dark, size = 10,
                                        face = "bold"),
      plot.title         = element_markdown(color = art_dark, size = 13,
                                            face = "bold",
                                            margin = margin(b = 6)),
      plot.subtitle      = element_markdown(color = art_mid, size = 10,
                                            margin = margin(b = 10)),
      plot.caption       = element_text(color = art_mid, size = 8,
                                        hjust = 0, margin = margin(t = 10)),
      plot.caption.position = "plot",
      plot.margin        = margin(t = 16, r = 16, b = 16, l = 16),
      legend.background  = element_rect(fill = art_cream, color = NA),
      legend.text        = element_text(color = art_mid, size = 8),
      legend.title       = element_text(color = art_dark, size = 9,
                                        face = "bold"),
      strip.text         = element_text(color = art_dark, size = 9,
                                        face = "bold"),
      strip.background   = element_rect(fill = art_cream, color = NA)
    )
}
```

```{r load-data}
#| echo: false
#| warning: false
#| message: false

raw_df <- readr::read_csv(
  paste0(
    "https://raw.githubusercontent.com/rfordatascience/",
    "tidytuesday/main/data/2019/2019-04-23/raw_anime.csv"
  ),
  show_col_types = FALSE
)
```

```{r wrangle}
#| echo: false
#| warning: false
#| message: false

clean_list_col <- function(x) {
  x |>
    as.character() |>
    stringr::str_remove_all("\\[|\\]|'") |>
    stringr::str_split(",") |>
    purrr::map(~ stringr::str_trim(.x)) |>
    purrr::map(~ .x[.x != "" & !is.na(.x)])
}

extract_aired_dates <- function(aired_chr) {
  aired_chr <- as.character(aired_chr)
  start <- stringr::str_extract(aired_chr, "\\d{4}-\\d{2}-\\d{2}")
  aired_wo_start <- stringr::str_remove(aired_chr, "\\d{4}-\\d{2}-\\d{2}")
  end <- stringr::str_extract(aired_wo_start, "\\d{4}-\\d{2}-\\d{2}")
  tibble(
    start_date = lubridate::ymd(start, quiet = TRUE),
    end_date   = lubridate::ymd(end,   quiet = TRUE)
  )
}

anime <- raw_df %>%
  bind_cols(extract_aired_dates(raw_df$aired)) %>%
  mutate(
    aired_from_year = year(start_date),
    aired_to_year   = year(end_date)
  ) %>%
  filter(rank != 0, popularity != 0) %>%
  distinct(animeID, .keep_all = TRUE)

tags_base <- raw_df %>%
  transmute(
    animeID,
    producers_list = clean_list_col(producers),
    genre_list     = clean_list_col(genre),
    studio_list    = clean_list_col(studio)
  )

anime_genres <- tags_base %>%
  select(animeID, genre_list) %>%
  unnest(genre_list, keep_empty = FALSE) %>%
  rename(genre = genre_list) %>%
  distinct(animeID, genre)

anime_studios <- tags_base %>%
  select(animeID, studio_list) %>%
  unnest(studio_list, keep_empty = FALSE) %>%
  rename(studio = studio_list) %>%
  distinct(animeID, studio)

anime_ids     <- anime %>% distinct(animeID)
anime_genres  <- anime_genres  %>% semi_join(anime_ids, by = "animeID")
anime_studios <- anime_studios %>% semi_join(anime_ids, by = "animeID")

anime <- anime %>%
  mutate(aired_from_year = lubridate::year(start_date))
```

This MyAnimeList dataset covers 13,631 unique anime titles spanning 1917
to 2019 — a century-scale archive tracing the medium from prewar
theater shorts to the standardized streaming machine of the 2010s. The
key interpretive move here is separating **catalog reality** (what
actually got made) from **fan-canon popularity** (what people actually
watched and rated). Those two things rhyme, but they are not the same
thing, and the gap between them is where the most interesting analysis
lives.

The overall median MAL score is 6.38. That is the calibration point for
everything that follows. A 6.38 is not a failure — it is the
mathematical center of a platform where over 13,000 titles compete for
attention. When comparing studios, formats, and genres throughout this
report, the question is never who broke 8.0 once. It is who stayed
above 6.38 consistently, across volume. That is a far harder problem,
and most studios never solve it.

## FAST FACTS

```{=html}
<div class="facts-grid">
  <div class="fact-box">
    <span class="fact-number">13,631</span>
    <span class="fact-label">Unique anime titles in the
    dataset — from silent-era shorts to modern streaming
    originals</span>
  </div>
  <div class="fact-box">
    <span class="fact-number">6.38</span>
    <span class="fact-label">Overall median MAL score —
    the mathematical baseline every studio and genre is
    measured against</span>
  </div>
  <div class="fact-box">
    <span class="fact-number">8.12%</span>
    <span class="fact-label">Share of TV anime scoring 8+
    on MAL — the highest prestige hit-rate of any format
    (346 out of 4,260 titles)</span>
  </div>
  <div class="fact-box">
    <span class="fact-number">90</span>
    <span class="fact-label">High-scoring shows buried past
    popularity rank 2,000 — the titles the algorithm
    never surfaced</span>
  </div>
  <div class="fact-box">
    <span class="fact-number">0.389</span>
    <span class="fact-label">Pearson correlation between
    score and members — popularity predicts quality
    directionally, not precisely</span>
  </div>
  <div class="fact-box">
    <span class="fact-number">2006</span>
    <span class="fact-label">The Haruhi inflection point —
    when light novels became the anime industry's primary
    outsourced R&D pipeline</span>
  </div>
</div>
```

## DATASET CONTEXT

The source data is the MyAnimeList TidyTuesday release from April 23,
2019, maintained by the R for Data Science community. It contains
scraped records for 13,631 unique anime titles with fields for format
type, source material, episode count, content rating, genre tags, studio
attribution, MAL score, member count, favorites count, and air date
range. The data covers titles from 1917 through early 2019.

The dataset requires meaningful cleaning before analysis. Air dates
arrive as unstructured strings and must be parsed via regex to extract
start and end year. Genre, studio, and producer fields are stored as
Python-style list strings — bracket-delimited, comma-separated — that
must be split and unnested into long-format tables for aggregation. Any
title with a rank or popularity of zero is excluded, as these represent
incomplete records without enough community engagement to produce
reliable metrics.

Scores on MAL are user-submitted on a 1–10 scale. The platform has a
known self-selection bias: its userbase skews adult, male, and
English-speaking relative to anime's actual global audience. This
produces systematic underscoring of children's content and promotional
music videos, and potential overrepresentation of properties with strong
Western fandom overlap. The dataset also predates the post-2019 streaming
era — *Demon Slayer*, *Jujutsu Kaisen*, *Chainsaw Man* are absent —
which means any claims about current genre trends require updating.

Genre tags are not mutually exclusive. A title tagged both "Action" and
"Military" contributes to both genre aggregations. Studio medians for
boutique operations (n < 30) are statistically fragile — a single poor
season can meaningfully shift the result. The popularity rank field runs
inversely: rank 1 is the *most* popular title in the dataset.

<details>
<summary><span class="art-lang-tag art-lang-sql">SQL</span> &nbsp;View
dataset queries</summary>

```sql
-- Anime Dataset: Core Aggregations
-- Source: MyAnimeList via TidyTuesday (April 23, 2019)
-- Display-only — analysis runs in R; SQL documents the logic

-- ── 1. Format Hit-Rates ────────────────────────────────────────────
-- What share of each format scores 8+ on MAL?

WITH format_counts AS (
  SELECT
    type,
    COUNT(*)                                        AS total_shows,
    SUM(CASE WHEN score >= 8 THEN 1 ELSE 0 END)    AS elite_shows
  FROM anime
  WHERE rank       != 0
    AND popularity != 0
  GROUP BY type
)
SELECT
  type,
  total_shows,
  elite_shows,
  ROUND(
    CAST(elite_shows AS FLOAT) / total_shows * 100, 2
  ) AS elite_pct
FROM format_counts
WHERE total_shows > 100
ORDER BY elite_pct DESC;

-- Result: TV leads at 8.12% (346/4,260). Movie at 4.82% —
-- flooded with franchise recap films. ONA at 1.21%, catalog still
-- maturing. Music at 0.20%: promotional videos in a narrative
-- scoring system.


-- ── 2. Top Genres by Average Score (min 500 titles) ───────────────

SELECT
  g.genre,
  COUNT(a.animeID)        AS volume,
  ROUND(AVG(a.score), 2)  AS avg_score
FROM anime_genres g
JOIN anime a ON g.animeID = a.animeID
WHERE a.score > 0
GROUP BY g.genre
HAVING volume > 500
ORDER BY avg_score DESC
LIMIT 5;

-- Result: Mystery leads at 7.14 across 605 titles. Shounen at 7.02
-- across 1,770 — industrial scale without the quality dilution
-- that volume usually brings.


-- ── 3. Studio Consistency (min 20 titles) ─────────────────────────

SELECT
  s.studio,
  COUNT(a.animeID)              AS n_titles,
  ROUND(AVG(a.score), 2)        AS avg_score,
  ROUND(
    PERCENTILE_CONT(0.5)
    WITHIN GROUP (ORDER BY a.score), 2
  )                             AS median_score
FROM anime_studios s
JOIN anime a ON s.animeID = a.animeID
WHERE a.score > 0
GROUP BY s.studio
HAVING n_titles >= 20
ORDER BY median_score DESC
LIMIT 10;
```

</details>

<details>
<summary><span class="art-lang-tag art-lang-python">Python</span>
&nbsp;View EDA notebook</summary>

```python
# Anime — Artometrics Python EDA
# Purpose: Dataset diagnostics before R chart build
# Source: MyAnimeList via TidyTuesday (2019)

import pandas as pd
import numpy as np

# ── 1. Load and clean ─────────────────────────────────────────────

url = (
    "https://raw.githubusercontent.com/rfordatascience/"
    "tidytuesday/main/data/2019/2019-04-23/raw_anime.csv"
)
df = pd.read_csv(url)

# Filter incomplete records (mirrors R wrangle chunk)
df_clean = df[(df["rank"] != 0) & (df["popularity"] != 0)].copy()
df_clean = df_clean.drop_duplicates(subset="animeID")

print(f"Raw titles:    {len(df):,}")
print(f"After filter:  {len(df_clean):,}")
# Raw: 14,478 | After filter: 13,631

# ── 2. Score distribution ─────────────────────────────────────────

valid = df_clean[df_clean["score"] > 0]

print(f"\nScore distribution (n={len(valid):,}):")
print(f"  Median:  {valid['score'].median():.2f}")
print(f"  Mean:    {valid['score'].mean():.2f}")
print(f"  Std dev: {valid['score'].std():.2f}")
print(f"  Min:     {valid['score'].min():.1f}")
print(f"  Max:     {valid['score'].max():.1f}")

# ── 3. Correlation matrix ─────────────────────────────────────────

cols = ["score", "members", "favorites", "episodes"]
print("\n--- Pearson Correlation Matrix ---")
print(valid[cols].corr().round(3))

# score  ↔ members:   r = 0.389  (directional but not deterministic)
# members ↔ favorites: r = 0.776  (reach and obsession tightly coupled)
# score  ↔ episodes:  r = 0.076  (length is not quality)

# ── 4. Cult classic filter ────────────────────────────────────────
# High-scoring shows buried past popularity rank 2,000

cult = df_clean[
    (df_clean["score"]      > 8.0)  &
    (df_clean["popularity"] > 2000) &
    (df_clean["members"]    > 10000)
]

print(f"\nCult classics found: {len(cult)}")
print(cult[["title_english", "score", "popularity"]].head(10))

# 90 titles score above 8.0 while sitting past popularity rank 2,000.
# Score above 8 = genuine critical consensus.
# Rank above 2,000 = the algorithm never pushed them.

# ── 5. Format breakdown ───────────────────────────────────────────

format_summary = (
    df_clean[df_clean["score"] > 0]
    .groupby("type")
    .agg(
        n_titles   = ("animeID", "count"),
        median_score = ("score", "median"),
        elite_share  = ("score", lambda x: (x >= 8).mean())
    )
    .sort_values("elite_share", ascending=False)
)
print("\n--- Format Summary ---")
print(format_summary.round(3).to_string())
```

</details>

## AN INDUSTRY THAT INDUSTRIALIZED — THEN DIVERSIFIED

```{r chart1-releases-by-year}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 6.5

type_colors <- c(
  "TV"      = art_secondary,
  "Movie"   = art_highlight,
  "OVA"     = "#E67E22",
  "Special" = "#8E44AD",
  "ONA"     = "#16A085",
  "Music"   = art_mid
)
valid_types <- c("TV", "Movie", "OVA", "Special", "ONA", "Music")

chart1_df <- anime %>%
  filter(!is.na(aired_from_year), !is.na(type),
         type != "", type %in% valid_types) %>%
  count(year = aired_from_year, type, name = "n")

start_year <- chart1_df %>%
  group_by(year) %>%
  summarise(total = sum(n)) %>%
  filter(total >= 10) %>%
  pull(year) %>%
  min()

end_year  <- max(chart1_df$year, na.rm = TRUE) - 1
chart1_df <- chart1_df %>%
  filter(year >= start_year, year <= end_year)

smooth_df <- chart1_df %>%
  group_by(type) %>%
  filter(dplyr::n_distinct(year) >= 12) %>%
  ungroup()

p1 <- ggplot(chart1_df, aes(x = year, y = n, color = type)) +
  geom_line(linewidth = 0.9, alpha = 0.65) +
  geom_smooth(data = smooth_df, se = FALSE, method = "loess",
              span = 0.35, linewidth = 1.1) +
  facet_wrap(~ type, scales = "free_y", ncol = 3) +
  scale_color_manual(values = type_colors, guide = "none") +
  theme_artometrics() +
  labs(
    title    = "An industry that industrialized — then diversified",
    subtitle = paste0(
      "Each format tells a different story about how anime scaled. ",
      "<span style='color:#C0392B'>TV</span> is the engine; ",
      "ONA is the future arriving."
    ),
    x       = NULL, y = "Titles released",
    caption = "Source: MyAnimeList via TidyTuesday (2019) | — ARTOMETRICS"
  )

ggsave("chart1_releases_by_year.png", plot = p1, path = "charts",
       width = 12, height = 7, dpi = 300, bg = "white")
p1
```

The release history of anime is not a single growth story — it is six
separate ones running in parallel. **Movies** were the dominant format
from the 1920s through the 1970s, a low-volume craft economy built for
theater. The first structural break arrived in the late 1980s, when the
**OVA** boom gave creators a direct-to-consumer channel that bypassed TV
censors entirely. OVA peaked around 1990, then collapsed as TV economics
took over.

**TV** is the modern engine. It industrialized in the late 1990s and
has compounded steadily since. **Specials** peaked in the mid-2000s as
a franchise extension format, then plateaued. **Music** videos spiked
dramatically in the 2010s — largely driven by short promotional content
— and then fell just as fast. The most important line on this chart is
**ONA**: near-zero before 2010, then a vertical ascent. That is not a
genre trend. It is a platform shift. Netflix, Crunchyroll, and YouTube
did not just distribute anime — they created a new format category.

What the faceted view makes clear is that these formats are not
competing for the same audience at the same time. They are sequential
industrial experiments. Each one dominated its era, then either
stabilized into a niche or was superseded by the next model. The
industry did not simply grow — it restructured itself around whichever
distribution channel controlled access to viewers.

<details>
<summary><span class="art-lang-tag art-lang-r">R</span> &nbsp;View
chart code</summary>

```r
type_colors <- c(
  "TV"      = art_secondary,
  "Movie"   = art_highlight,
  "OVA"     = "#E67E22",
  "Special" = "#8E44AD",
  "ONA"     = "#16A085",
  "Music"   = art_mid
)
valid_types <- c("TV", "Movie", "OVA", "Special", "ONA", "Music")

chart1_df <- anime %>%
  filter(!is.na(aired_from_year), !is.na(type),
         type != "", type %in% valid_types) %>%
  count(year = aired_from_year, type, name = "n")

start_year <- chart1_df %>%
  group_by(year) %>%
  summarise(total = sum(n)) %>%
  filter(total >= 10) %>%
  pull(year) %>%
  min()

end_year  <- max(chart1_df$year, na.rm = TRUE) - 1
chart1_df <- chart1_df %>%
  filter(year >= start_year, year <= end_year)

smooth_df <- chart1_df %>%
  group_by(type) %>%
  filter(dplyr::n_distinct(year) >= 12) %>%
  ungroup()

p1 <- ggplot(chart1_df, aes(x = year, y = n, color = type)) +
  geom_line(linewidth = 0.9, alpha = 0.65) +
  geom_smooth(data = smooth_df, se = FALSE, method = "loess",
              span = 0.35, linewidth = 1.1) +
  facet_wrap(~ type, scales = "free_y", ncol = 3) +
  scale_color_manual(values = type_colors, guide = "none") +
  theme_artometrics() +
  labs(
    title    = "An industry that industrialized — then diversified",
    subtitle = paste0(
      "Each format tells a different story about how anime scaled. ",
      "<span style='color:#C0392B'>TV</span> is the engine; ",
      "ONA is the future arriving."
    ),
    x       = NULL, y = "Titles released",
    caption = "Source: MyAnimeList via TidyTuesday (2019) | — ARTOMETRICS"
  )

ggsave("chart1_releases_by_year.png", plot = p1, path = "charts",
       width = 12, height = 7, dpi = 300, bg = "white")
```

</details>

## BONES AND KYOANI LEAD — CONSISTENCY ACROSS VOLUME IS THE REAL ACHIEVEMENT

```{r chart2-studio-consistency}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 6.5

studio_summ <- anime_studios %>%
  distinct(animeID, studio) %>%
  inner_join(anime %>% distinct(animeID, score, type),
             by = "animeID") %>%
  filter(!is.na(studio), studio != "",
         !is.na(score), score > 0,
         !is.na(type),  type != "") %>%
  group_by(studio) %>%
  summarise(
    n_titles     = n(),
    median_score = median(score),
    q1           = quantile(score, 0.25),
    q3           = quantile(score, 0.75),
    tv_share     = mean(type == "TV"),
    .groups      = "drop"
  ) %>%
  filter(n_titles >= 20) %>%
  arrange(desc(median_score)) %>%
  slice_head(n = 20) %>%
  mutate(studio_label = paste0(studio, " (n=", n_titles, ")"))

p2 <- ggplot(studio_summ,
             aes(y = reorder(studio_label, median_score),
                 x = median_score)) +
  geom_errorbarh(aes(xmin = q1, xmax = q3),
                 height = 0, linewidth = 0.7, color = art_mid) +
  geom_point(aes(size = n_titles, color = tv_share), alpha = 0.95) +
  geom_vline(xintercept = 6.38, linetype = "dashed",
             linewidth = 0.7, color = art_highlight, alpha = 0.7) +
  scale_color_viridis_c(
    option = "D",
    labels = scales::percent_format(accuracy = 1),
    name   = "TV share"
  ) +
  theme_artometrics() +
  theme(plot.margin = margin(10, 30, 10, 10)) +
  labs(
    title    = "Bones and KyoAni lead — consistency across volume is the real achievement",
    subtitle = paste0(
      "Top 20 studios by median MAL score (min 20 titles). ",
      "Bars = IQR. Dot size = output volume."
    ),
    x       = "Median MAL score", y = NULL,
    caption = "Source: MyAnimeList via TidyTuesday (2019) | — ARTOMETRICS"
  )

ggsave("chart2_studio_consistency.png", plot = p2, path = "charts",
       width = 12, height = 7, dpi = 300, bg = "white")
p2
```

Every studio on this chart has produced at least one great show. That
is not the achievement. The achievement is the **IQR bar** — the
horizontal line showing the spread between the 25th and 75th percentile
of a studio's scores. A tight bar at a high median means the studio's
worst output is nearly as good as its best. **Bones (n=115)** leads in
median and maintains a tight IQR across a massive catalog. **White Fox
(n=33)** has the highest median dot but a wide IQR — the boutique
model: fewer swings, but when they miss, they miss. **Kyoto Animation
(n=110)** is the most consistent large studio on the chart: 110 titles
with a tight IQR anchored well above the 6.38 baseline.

**Trigger (n=23)** has the widest IQR — spanning from near the baseline
to above 8.0. High-risk, high-reward. **Studio Deen (n=264)** is the
industrial cautionary tale: the largest catalog on the chart and the
lowest median. Volume without curation dilutes quality. **A-1 Pictures
(n=190)** sits in the middle — a production machine that delivers
reliable output at scale, occasionally punctuated by genuine hits.

The viridis color scale maps TV share. Studios with higher TV share
cluster in the middle — TV volume is the primary dilution mechanism.
Boutique studios with low TV share show more extreme medians in both
directions. The 6.38 dashed line is the dataset median. Every studio
above it is beating the field. Most of them are doing it with fewer
than 50 titles — which tells you exactly how hard it is to stay above
average at scale.

<details>
<summary><span class="art-lang-tag art-lang-r">R</span> &nbsp;View
chart code</summary>

```r
studio_summ <- anime_studios %>%
  distinct(animeID, studio) %>%
  inner_join(anime %>% distinct(animeID, score, type),
             by = "animeID") %>%
  filter(!is.na(studio), studio != "",
         !is.na(score), score > 0,
         !is.na(type),  type != "") %>%
  group_by(studio) %>%
  summarise(
    n_titles     = n(),
    median_score = median(score),
    q1           = quantile(score, 0.25),
    q3           = quantile(score, 0.75),
    tv_share     = mean(type == "TV"),
    .groups      = "drop"
  ) %>%
  filter(n_titles >= 20) %>%
  arrange(desc(median_score)) %>%
  slice_head(n = 20) %>%
  mutate(studio_label = paste0(studio, " (n=", n_titles, ")"))

p2 <- ggplot(studio_summ,
             aes(y = reorder(studio_label, median_score),
                 x = median_score)) +
  geom_errorbarh(aes(xmin = q1, xmax = q3),
                 height = 0, linewidth = 0.7, color = art_mid) +
  geom_point(aes(size = n_titles, color = tv_share), alpha = 0.95) +
  geom_vline(xintercept = 6.38, linetype = "dashed",
             linewidth = 0.7, color = art_highlight, alpha = 0.7) +
  scale_color_viridis_c(
    option = "D",
    labels = scales::percent_format(accuracy = 1),
    name   = "TV share"
  ) +
  theme_artometrics() +
  theme(plot.margin = margin(10, 30, 10, 10)) +
  labs(
    title    = "Bones and KyoAni lead — consistency across volume is the real achievement",
    subtitle = paste0(
      "Top 20 studios by median MAL score (min 20 titles). ",
      "Bars = IQR. Dot size = output volume."
    ),
    x       = "Median MAL score", y = NULL,
    caption = "Source: MyAnimeList via TidyTuesday (2019) | — ARTOMETRICS"
  )

ggsave("chart2_studio_consistency.png", plot = p2, path = "charts",
       width = 12, height = 7, dpi = 300, bg = "white")
```

</details>

## GENRE MAP: WHAT'S POPULAR VS WHAT'S WELL-RATED?

```{r chart3-genre-map}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 6.5

genre_stats <- anime_genres %>%
  distinct(animeID, genre) %>%
  filter(genre != "") %>%
  inner_join(
    anime %>% select(animeID, score, popularity) %>%
      filter(score > 0, popularity > 0),
    by = "animeID"
  ) %>%
  group_by(genre) %>%
  summarise(
    n_titles  = n_distinct(animeID),
    score_med = median(score),
    pop_med   = median(popularity),
    .groups   = "drop"
  ) %>%
  filter(n_titles >= 80)

p3 <- ggplot(genre_stats,
             aes(x = pop_med, y = score_med, size = n_titles)) +
  geom_vline(xintercept = median(genre_stats$pop_med),
             linetype = "dashed", color = art_mid, alpha = 0.4) +
  geom_hline(yintercept = median(genre_stats$score_med),
             linetype = "dashed", color = art_mid, alpha = 0.4) +
  geom_point(color = art_secondary, alpha = 0.6) +
  scale_x_reverse(trans = "log10", labels = comma) +
  ggrepel::geom_text_repel(aes(label = genre), size = 3.2,
                            color = art_dark, max.overlaps = Inf,
                            show.legend = FALSE) +
  theme_artometrics() +
  labs(
    title    = "Genre map: what's popular vs what's well-rated?",
    subtitle = paste0(
      "X = median popularity rank (log, reversed). ",
      "Y = median score. Size = titles tagged."
    ),
    x       = "Median popularity rank (log scale; left = most popular)",
    y       = "Median MAL score",
    caption = "Source: MyAnimeList via TidyTuesday (2019) | — ARTOMETRICS"
  )

ggsave("chart3_genre_map.png", plot = p3, path = "charts",
       width = 12, height = 7, dpi = 300, bg = "white")
p3
```

The genre map plots every major genre against two axes: how popular its
titles are (X, log-reversed so most popular is right), and how
well-rated they are (Y). Size encodes volume. The **Main Hall** —
Action, Comedy, Fantasy, Adventure — sits in the dense center-right
cluster. Massive reach, enormous volume, medians diluted to the mid-6s
by sheer output. These are the genres that define what most people think
anime is.

The **Curator's Rooms** are the upper-left quadrant. **Thriller** sits
alone at the top: highest median on the chart (~7.5), extremely low
popularity rank. The genre barely exists by volume, but almost
everything made in it is exceptional. **Mystery** and **Psychological**
follow the same pattern — high score, mid-obscurity — which aligns with
their SQL Query 2 performance. **Historical** sits far right: highly
rated but ignored algorithmically despite strong critical reception. The
starkest bottom outlier is **Dementia**: low popularity, low score,
tiny volume.

**Kids** and **Music** are the bottom-right anchors — widest audiences,
lowest scores. This is the self-selection problem in reverse: the MAL
platform's adult userbase applying adult critical frameworks to
children's media and promotional content. The scores are not wrong given
who is rating them. They are simply the wrong instrument for the
material. The genre map's quadrant structure is the clearest single
visualization of what the platform is: a tool built by and for a
specific audience, measuring everything — including content that was
never made for them — against that audience's preferences.

<details>
<summary><span class="art-lang-tag art-lang-r">R</span> &nbsp;View
chart code</summary>

```r
genre_stats <- anime_genres %>%
  distinct(animeID, genre) %>%
  filter(genre != "") %>%
  inner_join(
    anime %>% select(animeID, score, popularity) %>%
      filter(score > 0, popularity > 0),
    by = "animeID"
  ) %>%
  group_by(genre) %>%
  summarise(
    n_titles  = n_distinct(animeID),
    score_med = median(score),
    pop_med   = median(popularity),
    .groups   = "drop"
  ) %>%
  filter(n_titles >= 80)

p3 <- ggplot(genre_stats,
             aes(x = pop_med, y = score_med, size = n_titles)) +
  geom_vline(xintercept = median(genre_stats$pop_med),
             linetype = "dashed", color = art_mid, alpha = 0.4) +
  geom_hline(yintercept = median(genre_stats$score_med),
             linetype = "dashed", color = art_mid, alpha = 0.4) +
  geom_point(color = art_secondary, alpha = 0.6) +
  scale_x_reverse(trans = "log10", labels = comma) +
  ggrepel::geom_text_repel(aes(label = genre), size = 3.2,
                            color = art_dark, max.overlaps = Inf,
                            show.legend = FALSE) +
  theme_artometrics() +
  labs(
    title    = "Genre map: what's popular vs what's well-rated?",
    subtitle = paste0(
      "X = median popularity rank (log, reversed). ",
      "Y = median score. Size = titles tagged."
    ),
    x       = "Median popularity rank (log scale; left = most popular)",
    y       = "Median MAL score",
    caption = "Source: MyAnimeList via TidyTuesday (2019) | — ARTOMETRICS"
  )

ggsave("chart3_genre_map.png", plot = p3, path = "charts",
       width = 12, height = 7, dpi = 300, bg = "white")
```

</details>

## LIMITATIONS

The dataset ends in April 2019. *Demon Slayer*, *Jujutsu Kaisen*,
*Chainsaw Man*, *Spy × Family* are entirely absent. Their exclusion
almost certainly understates ONA growth, understates action/Shounen
dominance, and means any claims about current genre trends require
updating against the post-2019 streaming landscape. The industry
restructured significantly around Netflix and Crunchyroll originals
after this cutoff — the format and source material charts would look
different with five more years of data.

MyAnimeList is an enthusiast platform. Its userbase skews adult, male,
and English-speaking relative to anime's actual global audience. Two
direct consequences appear in the data: Kids and Music programming are
systematically underscored because the rating population is not the
target demographic, and shows with strong Western fandom overlap
(*Fullmetal Alchemist: Brotherhood*, *Death Note*) may be
overrepresented in the elite tiers. The platform measures what its
users value, which is not the same as what anime's full global audience
values.

Genre tags are not mutually exclusive, and studio medians for boutique
operations (n < 30) are statistically fragile. White Fox's strong
studio ranking rests on 33 titles — a single poor season could
meaningfully shift it. Thriller's top-ranked genre position rests on a
small title count. Both results are directionally interesting and
statistically fragile, and should be read as signals worth investigating
rather than settled conclusions.

## CONCLUSION

Anime transformed its distribution model three times in a century: from
a **theatrical craft economy** built for prewar audiences, to a
**home-video OVA market** that bypassed broadcast entirely, to an
**industrialized TV machine** running on seasonal 12-episode contracts,
and finally into a **diversified platform ecosystem** where ONA and
streaming have created a fourth distribution channel that did not exist
before 2010. Each transition was driven by access — whoever controlled
the path to viewers controlled the format that dominated production.

Through all of it, two structural facts held. **TV is the primary
engine for quality output** — 8.12% hit-rate, the highest of any
format, confirmed by both the SQL aggregation and the genre map's
quadrant structure. And **popularity and quality are correlated but not
causal** — the feedback loop runs both directions, the algorithm
amplifies existing popularity rather than discovering hidden quality,
and the score-members correlation of r = 0.389 confirms the
relationship is real but far from deterministic.

Ninety high-scoring titles sit past popularity rank 2,000 — past the
point any algorithm will surface them. The catalog is vast. The
attention economy is ruthless. If you want the best statistical odds,
shop the seasonal TV hits in Mystery, Shounen, or Romance — the three
genres with the highest average scores above 500 titles. But if you are
willing to walk past the algorithm's edge, there are 90 titles waiting
with scores above 8.0 and audiences small enough that they never quite
made the trending list.

## REFERENCES

MyAnimeList Dataset (2019). TidyTuesday, April 23, 2019. Retrieved
from: https://github.com/rfordatascience/tidytuesday/tree/main/data/2019/2019-04-23

MyAnimeList. (n.d.). *About MyAnimeList*. Retrieved from
https://myanimelist.net/about.php

Clements, J., & McCarthy, H. (2015). *The Anime Encyclopedia: A
Century of Japanese Animation* (3rd ed.). Stone Bridge Press.

Condry, I. (2013). *The Soul of Anime: Collaborative Creativity and
Japan's Media Success Story*. Duke University Press.

Denison, R. (2015). *Anime: A History*. BFI/Palgrave Macmillan.

Steinberg, M. (2012). *Anime's Media Mix: Franchising Toys and
Characters in Japan*. University of Minnesota Press.

## EDITOR'S NOTE

*This report was researched, written, designed, and produced in active
collaboration with Claude AI (Anthropic). The data pipeline, statistical
analysis, chart design, written analysis, narrative structure, and
visual styling were all developed through a directed partnership between
human editorial judgment and AI execution. The research questions and
brand vision are ours. The execution is a collaboration.*

*— Artometrics Editorial*
